Plug-and-play dual-tree algorithm runtime analysis

### 2014-11-18 bounds-files/sec03-approx-kde.tex
Some further refactoring.

A tree-independent algorithm for solving approximate kernel density estimation
is given as a \texttt{BaseCase()} function in Algorithm \ref{alg:kde_base_case}
and a \texttt{Score()} function in Algorithm \ref{alg:kde_score} (a correctness
proof can be found in \cite{curtin2013tree}).  This function
assumes a shift-invariant kernel $\mathcal{K}(p_i, p_j) = \mathcal{K}(d( p_i,
p_j ))$.  The list $f_p$ holds partial kernel density estimates for
each query point, and the list $f_n$ holds partial kernel density
estimates for each query node.  At the beginning of the dual-tree traversal, the
lists $f_p$ and $f_n$, which are both of size $O(N)$, are each
initialized to 0.

@@

Note that this version is far simpler than the algorithm proposed in,
e.g., \cite{gray2003nonparametric}; however, it is sufficient for our runtime
analysis.  Real-world implementations, such as the one found in \textbf{mlpack}
\cite{mlpack2013}, tend to be far more complex.

@@

  \caption{Approximate kernel density estimation \texttt{Score()}}
  \label{alg:kde_score}
\end{algorithm}

\begin{thm}
Assume that $\mathcal{K}(p_i, p_j) = \mathcal{K}\left(d(p_i, p_j)\right)$ is a shift-invariant kernel with some bandwidth $h > 0$, where $\mathcal{K}(x)$ is smooth, non-negative, monotonically decreasing, and concave for $x \in [0, h]$ and convex for $x \in (h, \infty)$. Then, given a query set $S_q$ and a reference set $S_r$ with
expansion constant $c_r$, an inverse degree of bichromaticity $\nu$ and using the
approximate kernel density estimation \texttt{BaseCase()} and \texttt{Score()}
as given in Algorithms \ref{alg:kde_base_case} and \ref{alg:kde_score},
respectively, with the traversal given in Algorithm \ref{alg:cover-tree-dual},
the running time of approximate kernel density estimation for some $\epsilon$
error parameter is bounded by
%\begin{equation}
$O(c_r^{8 + \log_2 \zeta} \nu N)$
%\end{equation}
where $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$.

@@

\begin{proof}
It is clear that \texttt{BaseCase()} and \texttt{Score()} both take $O(1)$ time,
so Theorem \ref{thm:ct-runtime} implies the total runtime of the dual-tree
algorithm is bounded by $O(c_r^4 \nu |R^*| N)$.  Thus, we will bound $|R^*|$
using techniques related to those used in \cite{ram2009}.  The bounding of
$|R^*|$ is split into two sections: first, show that when the scale $s_r^{\max}$
is small enough, $R^*$ is empty.  Second, bound $R^*$ when $s_r^{\max}$ is larger.

@@

For the Gaussian kernel with bandwidth $\sigma$ defined by $\mathcal{K}_g(d) =
\exp(-d^2 / (2 \sigma^2))$, $\zeta$ does not
depend on the kernel bandwidth; only
the approximation parameter $\epsilon$.  For this kernel, $h = \sigma$ and
therefore $-\mathcal{K}'_g(h) = \sigma^{-1} e^{-1 / 2}$.  Additionally,
$\mathcal{K}_g^{-1}(\epsilon) = \sigma \sqrt{2 \ln (1 / \epsilon)}$.  This means
that for the Gaussian kernel, $\zeta = \sqrt{(-2 \ln \epsilon) / (e
\epsilon^2)}$.  Again, as $\epsilon \to 0$, the
runtime diverges; however, note that there is no dependence on the kernel
bandwidth $\sigma$.  For a reasonably chosen $\epsilon = 0.05$, the runtime is
approximately $O(c_r^{8.89} \nu N)$; for $\epsilon = 0.01$, the runtime is
approximately $O(c_r^{11.52} \nu N)$.  For very small $\epsilon = 0.00001$, the
runtime is approximately $O(c_r^{22.15} \nu N)$.

@@

>>>>

A tree-independent algorithm for solving approximate kernel density estimation
with absolute value approximation
is given as a \texttt{BaseCase()} function in Algorithm \ref{alg:kde_base_case}
and a \texttt{Score()} function in Algorithm \ref{alg:kde_score}  \citep[a
correctness proof can be found in][]{curtin2013tree}.  This function
assumes a shift-invariant kernel $\mathcal{K}(p_i, p_j) = \mathcal{K}(d( p_i,
p_j ))$.  The list $f_p$ holds partial kernel density estimates for
each query point, and the list $f_n$ holds partial kernel density
estimates for each query node.  At the beginning of the dual-tree traversal, the
lists $f_p$ and $f_n$, which are both of size $O(N)$, are each
initialized to 0.

@@

Note that this version is far simpler than other dual-tree algorithms that have
been proposed for approximate kernel density estimation \citep[see, for
instance][]{gray2003nonparametric}; however, this version is sufficient for our
runtime analysis.  Real-world implementations, such as the one found in
\textbf{mlpack} \citep{mlpack2013}, tend to be far more complex.

@@

  \caption{Absolute-value approximate kernel density estimation
\texttt{Score()}}
  \label{alg:kde_score}
\end{algorithm}

    \IF{$\mathcal{K}(d_{\min}(\mathscr{N}_q, \mathscr{N}_r)) -
\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r)) /
\mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r)) < \epsilon$}
      \STATE $f_n(\mathscr{N}_q) \gets f_n(\mathscr{N}_q) + |
\mathscr{D}^p(\mathscr{N}_r) | \left(\mathcal{K}(d_{\min}(\mathscr{N}_q,
\mathscr{N}_r)) + \mathcal{K}(d_{\max}(\mathscr{N}_q, \mathscr{N}_r))\right) /\;2$
      \RETURN $\infty$
    \ENDIF

  \caption{Relative-value approximate kernel density estimation
\texttt{Score()}}
  \label{alg:kde_rva_score}
\end{algorithm}

\begin{thm}
Assume that $\mathcal{K}(p_i, p_j) = \mathcal{K}\left(d(p_i, p_j)\right)$ is a
shift-invariant kernel with some bandwidth $h > 0$, where $\mathcal{K}(x)$ is
smooth, non-negative, monotonically decreasing, and concave for $x \in [0, h]$
and convex for $x \in (h, \infty)$. Then, given a query set $S_q$ and a
reference set $S_r$ with
expansion constant $c_r$, and using the
approximate kernel density estimation \texttt{BaseCase()} and \texttt{Score()}
as given in Algorithms \ref{alg:kde_base_case} and \ref{alg:kde_score},
respectively, with the traversal given in Algorithm \ref{alg:cover-tree-dual},
the running time of approximate kernel density estimation for some $\epsilon$
error parameter is bounded by
%\begin{equation}
$O(c_r^{8 + \log_2 \zeta} (N + \theta))$
%\end{equation}
where $\zeta = -\mathcal{K}'(h) \mathcal{K}^{-1}(\epsilon) \epsilon^{-1}$ and
$\theta$ as in Theorem \ref{thm:ct-runtime}.

@@

\begin{proof}
It is clear that \texttt{BaseCase()} and \texttt{Score()} both take $O(1)$ time,
so Theorem \ref{thm:ct-runtime} implies the total runtime of the dual-tree
algorithm is bounded by $O(c_r^4 |R^*| (N + \theta))$.  Thus, we will bound
$|R^*|$ using techniques related to those used by \citet{ram2009}.  The bounding
of $|R^*|$ is split into two sections: first, show that when the scale
$s_r^{\max}$ is small enough, $R^*$ is empty.  Second, bound $R^*$ when
$s_r^{\max}$ is larger.

@@

For the Gaussian kernel with bandwidth $\sigma$ defined by $\mathcal{K}_g(d) =
\exp(-d^2 / (2 \sigma^2))$, $\zeta$ does not
depend on the kernel bandwidth; only
the approximation parameter $\epsilon$.  For this kernel, $h = \sigma$ and
therefore $-\mathcal{K}'_g(h) = \sigma^{-1} e^{-1 / 2}$.  Additionally,
$\mathcal{K}_g^{-1}(\epsilon) = \sigma \sqrt{2 \ln (1 / \epsilon)}$.  This means
that for the Gaussian kernel, $\zeta = \sqrt{(-2 \ln \epsilon) / (e
\epsilon^2)}$.  Again, as $\epsilon \to 0$, the
runtime diverges; however, note that there is no dependence on the kernel
bandwidth $\sigma$.  For a reasonably chosen $\epsilon = 0.05$, the runtime is
approximately $O(c_r^{8.89} (N + \theta))$; for $\epsilon = 0.01$, the runtime
is approximately $O(c_r^{11.52} (N + \theta))$.  For very small $\epsilon =
0.00001$, the runtime is approximately $O(c_r^{22.15} (N + \theta))$.

@@

\subsection{Relative Value Approximation}

Approximate kernel density estimation using relative-value approximation may be
bounded in a similar manner.  A \texttt{Score()} function is given in Algorithm
\ref{alg:kde_rva_score}.

### 2014-11-18 bounds-files/sec04-range-search.tex
Remove negative vspace and remove NIPS formatting.
Some further refactoring.

In the range search problem, the task is to find the set of reference points
%\begin{equation}
$N[p_q] = \{ p_r \in S_r : d(p_q, p_r) \in [l, u] \}$
%\end{equation}
for each query point $p_q$, where $[l, u]$ is the given
range.  The range count problem is practically identical, but only the size of
the set, $|N[p_q]|$, is desired.
Our proof works for both of these algorithms
identically, but we will focus on range search.  A \texttt{BaseCase()} and
\texttt{Score()} function are given in Algorithms \ref{alg:rs_bc} and
\ref{alg:rs_sc}, respectively (a correctness proof can be found in
\cite{curtin2013tree}).  The sets $N[p_q]$ (for each $p_q$) are
initialized to $\emptyset$ at the beginning of the traversal.

@@

\vspace*{-1em}
\begin{equation}
\left| \{ p_r: p_r \in S_r, (1 - \alpha) l \le d(p_q, p_r) \le (1 + \alpha) u \}
\right| = O(1),
\end{equation}
\vspace*{-1.5em}

\noindent the running time of range search or range count is bounded by
$O(c_r^{8 + \beta} \nu N)$ with
%\begin{equation}
$\beta = \log_2 (1 + \alpha^{-1})$.
%\end{equation}
\label{thm:rs}
\end{thm}

\vspace*{-2.5em}
\begin{proof}
Both \texttt{BaseCase()} (Algorithm \ref{alg:rs_bc}) and \texttt{Score()}
(Algorithm \ref{alg:rs_sc}) take $O(1)$ time.  Therefore, using Lemma
\ref{thm:ct-runtime} we know that the runtime of the algorithm is bounded by
$O(c_r^4 \nu |R^*| N)$. %  Thus, our only task is to bound the maximum size of the
%reference set, $|R^*|$.
By the pruning rule, for a query node
$\mathscr{N}_q$, the reference set $R^*$
is made up of reference nodes $\mathscr{N}_r$ that are within a margin of
$2^{s_q + 1} + 2^{s_r + 1} \le 2^{s_r^{\max} + 2}$ of the range $[l, u]$.  Given
that $p_r$ is the point in $\mathscr{N}_r$,

\vspace*{-1em}
\begin{equation}
%p_r &\in& \left( B_{S_r}(p_q, u + \lambda_q + \lambda_r) \cap
%C_{s_r^{\max}}\right) \setminus \left( B_{S_r}(p_q, l - \lambda_q - \lambda_r)
%\cap C_{s_r^{\max}} \right) \\
p_r \in \left( B_{S_r}(p_q, u + 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}}\right)\setminus \left( B_{S_r}(p_q, l - 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}} \right). \label{eqn:rsballs}
\end{equation}
\vspace*{-1em}

@@

%\vspace*{-1em}
%\begin{eqnarray}
%|B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| &\le& |B_{S_r}(p_r, (1 +
%\alpha^{-1}) 2^{s_r^{\max} + 3})| \\
%|B_{S_r}(p_q, u + 2^{s_r^{\max} + 1})| - |B_{S_r}(p_q, l - 2^{s_r^{\max} + 1})|
%&\le& |B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})| \nonumber \\
%&\le& |B_{S_r}(p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| \nonumber \\
%&\le& |B_{S_r \cup \{ p_q \}}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})
%\nonumber \\
%&\le& c_{qr}^{(2 + \log_2(1 + \alpha^{-1}))} |B_{S_r \cup \{ p_q \}}(p_q, 2^{s_r^{\max} -
%1})| \nonumber \\
%&\le& c_r^{(4 + \log_2(1 + \alpha^{-1}))} | B_{S_r}(p_r, 2^{s_r^{\max} - 1}) |.
%&\le& (1 + \alpha^{-1}) c_r^{3 + \log_2 c_r} | B_{S_r}(p_r, 2^{s_r^{\max} - 1})
%|.
%&\le& (1 + \alpha^{-1}) c_{qr}^{2 + \log_2 c_{qr}}
%\end{eqnarray}
%\vspace*{-1em}

@@

%\vspace*{-1em}
%\begin{equation}
%\left| B_{S_r}\left(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2}\right) \cap C_{s_r^{\max}}
%\right| \le \frac{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} +
%3}\right) \right|}{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} -
%1}\right) \right|} \le c_r^{4 + \log_2 (1 + \alpha^{-1})}.
%\end{equation}
%\vspace*{-1em}

@@

\vspace*{-1em}
\begin{equation}
B_{S_r}(p_q, u + 2^{s_r^{\max} + 1}) \setminus B_{S_r}(p_q, l - 2^{s_r^{\max} +
1}) \subseteq B_{S_r}(p_q, (1 + \alpha) u) \setminus B_{S_r}(p_q, (1 - \alpha)
l)
\end{equation}

\noindent and, due to the assumptions, $| B_{S_r}(p_q, (1 + \alpha) u) | - |
B_{S_r}(p_q, (1 - \alpha) l) | = O(1)$, $|R^*| = O(1)$.  Thus, the theorem
holds.
\end{proof}
\vspace*{-0.4em}

@@

To understand the relation of the running time to the bounding quantity
$\alpha$, consider (for the monochromatic case with $\nu \sim O(1)$) $\alpha =
(1 / 3)$; this yields a running time of $O(c^8 N)$.  $\alpha = (1 / 7)$ yields
$O(c^9 N)$, $\alpha = (1 / 15)$ yields $O(c^{10} N)$, and so forth.  As $\alpha$
gets smaller, the exponent on $c$ gets larger, and diverges as $\alpha \to 0$.

@@

Alternately, if the assumption is not valid but it can be shown that the number
of points in the range $[(1 - \alpha)l, (1 + \alpha)u]$ is $O(k)$ where $k$ has
some dependence on $N$, then the running time can be easily bounded as $O(c^{4 +
\beta} kN)$, in the monochromatic case with $\nu \sim O(1)$.

>>>>

In the range search problem, the task is to find the set of reference points
%\begin{equation}
$N[p_q] = \{ p_r \in S_r : d(p_q, p_r) \in [l, u] \}$
%\end{equation}
for each query point $p_q$, where $[l, u]$ is the given
range.  The range count problem is practically identical, but only the size of
the set, $|N[p_q]|$, is desired.
Our proof works for both of these algorithms
identically, but we will focus on range search.  A \texttt{BaseCase()} and
\texttt{Score()} function are given in Algorithms \ref{alg:rs_bc} and
\ref{alg:rs_sc}, respectively \citep[a correctness proof can be found
in][]{curtin2013tree}.  The sets $N[p_q]$ (for each $p_q$) are
initialized to $\emptyset$ at the beginning of the traversal.

@@

\begin{equation}
\left| \{ p_r: p_r \in S_r, (1 - \alpha) l \le d(p_q, p_r) \le (1 + \alpha) u \}
\right| = O(1),
\end{equation}

\noindent the running time of range search or range count is bounded by
$O(c_r^{8 + \beta} (N + \theta))$ with $\theta$ defined as in Theorem
\ref{thm:ct-runtime} and
%\begin{equation}
$\beta = \log_2 (1 + \alpha^{-1})$.
%\end{equation}
\label{thm:rs}
\end{thm}

\begin{proof}
Both \texttt{BaseCase()} (Algorithm \ref{alg:rs_bc}) and \texttt{Score()}
(Algorithm \ref{alg:rs_sc}) take $O(1)$ time.  Therefore, using Lemma
\ref{thm:ct-runtime} we know that the runtime of the algorithm is bounded by
$O(c_r^4 |R^*| (N + \theta))$. %  Thus, our only task is to bound the maximum size of the
%reference set, $|R^*|$.
By the pruning rule, for a query node
$\mathscr{N}_q$, the reference set $R^*$
is made up of reference nodes $\mathscr{N}_r$ that are within a margin of
$2^{s_q + 1} + 2^{s_r + 1} \le 2^{s_r^{\max} + 2}$ of the range $[l, u]$.  Given
that $p_r$ is the point in $\mathscr{N}_r$,

\begin{equation}
%p_r &\in& \left( B_{S_r}(p_q, u + \lambda_q + \lambda_r) \cap
%C_{s_r^{\max}}\right) \setminus \left( B_{S_r}(p_q, l - \lambda_q - \lambda_r)
%\cap C_{s_r^{\max}} \right) \\
p_r \in \left( B_{S_r}(p_q, u + 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}}\right)\setminus \left( B_{S_r}(p_q, l - 2^{s_r^{\max} + 2}) \cap
C_{s_r^{\max}} \right). \label{eqn:rsballs}
\end{equation}

@@

%\begin{eqnarray}
%|B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| &\le& |B_{S_r}(p_r, (1 +
%\alpha^{-1}) 2^{s_r^{\max} + 3})| \\
%|B_{S_r}(p_q, u + 2^{s_r^{\max} + 1})| - |B_{S_r}(p_q, l - 2^{s_r^{\max} + 1})|
%&\le& |B_{S_r}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})| \nonumber \\
%&\le& |B_{S_r}(p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2})| \nonumber \\
%&\le& |B_{S_r \cup \{ p_q \}}(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 1})
%\nonumber \\
%&\le& c_{qr}^{(2 + \log_2(1 + \alpha^{-1}))} |B_{S_r \cup \{ p_q \}}(p_q, 2^{s_r^{\max} -
%1})| \nonumber \\
%&\le& c_r^{(4 + \log_2(1 + \alpha^{-1}))} | B_{S_r}(p_r, 2^{s_r^{\max} - 1}) |.
%&\le& (1 + \alpha^{-1}) c_r^{3 + \log_2 c_r} | B_{S_r}(p_r, 2^{s_r^{\max} - 1})
%|.
%&\le& (1 + \alpha^{-1}) c_{qr}^{2 + \log_2 c_{qr}}
%\end{eqnarray}

@@

%\begin{equation}
%\left| B_{S_r}\left(p_q, (1 + \alpha^{-1}) 2^{s_r^{\max} + 2}\right) \cap C_{s_r^{\max}}
%\right| \le \frac{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} +
%3}\right) \right|}{\left| B_{S_r}\left( p_r, (1 + \alpha^{-1}) 2^{s_r^{\max} -
%1}\right) \right|} \le c_r^{4 + \log_2 (1 + \alpha^{-1})}.
%\end{equation}

@@

\begin{equation}
B_{S_r}(p_q, u + 2^{s_r^{\max} + 1}) \setminus B_{S_r}(p_q, l - 2^{s_r^{\max} +
1}) \subseteq B_{S_r}(p_q, (1 + \alpha) u) \setminus B_{S_r}(p_q, (1 - \alpha)
l)
\end{equation}

\noindent and, due to the assumptions, $| B_{S_r}(p_q, (1 + \alpha) u) | - |
B_{S_r}(p_q, (1 - \alpha) l) | = O(1)$, $|R^*| = O(1)$.  Thus, the theorem
holds.
\end{proof}

@@

To understand the relation of the running time to the bounding quantity
$\alpha$, consider $\alpha =
(1 / 3)$; this yields a running time of $O(c^8 (N + \theta))$.  $\alpha = (1 /
7)$ yields $O(c^9 (N + \theta))$, $\alpha = (1 / 15)$ yields $O(c^{10} (N +
\theta))$, and so forth.  As $\alpha$ gets smaller, the exponent on $c$ gets
larger, and diverges as $\alpha \to 0$.

@@

Alternately, if the assumption is not valid but it can be shown that the number
of points in the range $[(1 - \alpha)l, (1 + \alpha)u]$ is $O(k)$ where $k$ has
some dependence on $N$, then the running time can be easily bounded as $O(c^{4 +
\beta} k(N + \theta))$.

### 2014-11-18 bounds.tex
mathtools for equation alignment.
Some further refactoring.

\documentclass[twoside,11pt]{article} % For LaTeX2e
\usepackage{jmlr2e}
\usepackage{hyperref}
\hypersetup{colorlinks,
            urlcolor=blue,
            citecolor=green,
            linkcolor=red}
\usepackage{url}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
\usepackage{mathrsfs}
%\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{color}

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.  We
then investigate the empirical performance of cover trees for these algorithms,
with a focus on understanding the dataset-dependent constants that our bounds
use.

@@

\section{Tree Imbalance}
\label{sec:imbalance}

@@

\section{Range Search and Range Count}
\label{sec:rs}

@@

>>>>

\documentclass[twoside,11pt]{article} % For LaTeX2e
\usepackage{jmlr2e}
\usepackage{hyperref}
\hypersetup{colorlinks,
            urlcolor=blue,
            citecolor=green,
            linkcolor=red}
\usepackage{url}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
\usepackage{mathrsfs}
%\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{color}
\usepackage{mathtools}

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.  We
then investigate the empirical performance of cover trees for these algorithms
in Section \ref{sec:emp}, with a focus on understanding the dataset-dependent
constants that our bounds use.

@@

\section{Tree imbalance}
\label{sec:imbalance}

@@

\section{Range search and range count}
\label{sec:rs}

@@

\section{Empirical investigations}
\label{sec:emp}

\input{bounds-files/empirical.tex}
